Imputation of Incomplete Multilevel Data with \pkg{mice}

### 2022-06-23 Manuscript/Imputation_of_Incomplete_Multilevel_Data.Rmd
Add forest plot for popularity data

<!-- $$ -->
<!-- \texttt{popular} \sim  \texttt{1 + sex + texp + sex:texp + (1 + sex | school)} -->
<!-- $$ -->
```{r}
mod <- popular ~ 1 + sex + texp + sex:texp + (1 + sex | school)
```

```{r include=FALSE}
# data
popcomp <- foreign::read.spss("../Data/popular.sav", to.data.frame = TRUE, use.value.labels = FALSE) %>% 
  mutate(school = SCHOOL,
         popular = POPULAR,
         sex = SEX,
         texp = TEXP,
         .keep = "none")

est_true <- lme4::lmer(mod, popcomp) %>% broom.mixed::tidy()
# testEstimates(as.mitml.result(fit_ignored), var.comp = TRUE)

popmis <- mice::popmis[, c("school", "popular", "sex", "texp")] 
```

@@

```{r}
popmis <- popmis[, c("school", "popular", "sex", "texp")] 
```

@@

This shows us that both `sex` and `texp` may be useful imputation model predictors. Moreover, the missingness in `popular` may depend on the observed values of other variables. We'll highlight one other variable to illustrate, but ideally one would inspect all relations. The questions we'll ask are: 'Does the missing data of pupil popularity (`popular`) depend on observed teacher popularity (`texp`)?'. This can be evaluated statistically, but visual inspection usually suffices. We'll make a histogram of `texp` separately for the pupils with known popularity and missing popularity.

Plot the histogram for teacher experience conditional on the missingness indicator of `popular`:

```{r pop-hist}
ggmice(popmis, aes(texp)) +
  geom_histogram(fill = "white") +
  facet_grid(. ~ is.na(popular), scales = "free", labeller = label_both)
```

This shows us that there are no apparent differences in the distribution of `texp` depending on the missingness indicator of `popular` (t = `r t.test(popmis$texp ~ is.na(popmis$popular)) %>% broom::tidy(.) %>% .[, c("statistic", "p.value")] %>% round(., 3) %>% unlist() %>% paste(collapse = ",  p = ")`).
 [TODO: think about what is a meaningful rule of thumb to signal that the user should be worried?]

@@

Analyze the imputations:
```{r pop-ignored-fit, echo=TRUE, message=FALSE, warning=FALSE}
fit <- with(imp, 
            lmer(popular ~ 1 + sex + texp + sex:texp + (1 + sex | school))) 
```

@@

Analyze the imputations:
```{r, echo=TRUE, message=FALSE, warning=FALSE}
fit <- with(imp, 
            lmer(popular ~ 1 + sex + texp + sex:texp + (1 + sex | school))) 
```

>>>>

<!-- $$ -->
<!-- \texttt{popular} \sim  \texttt{1 + sex + texp + sex:texp + (1 + sex | school)} -->
<!-- $$ -->
```{r}
mod <- popular ~ 1 + sex + (1 | school)
```

```{r include=FALSE}
# data
popcomp <- foreign::read.spss("../Data/popular.sav", to.data.frame = TRUE, use.value.labels = FALSE) %>% 
  mutate(school = SCHOOL,
         popular = POPULAR,
         sex = SEX,
         .keep = "none")

est_true <- lme4::lmer(mod, popcomp) %>% broom.mixed::tidy()
# testEstimates(as.mitml.result(fit_ignored), var.comp = TRUE)
```

The true effect is:

```{r echo=FALSE}
fits <-
  purrr::map_dfr(
    unique(popcomp$school),
    ~ {
      lm(
        mod,
        data = filter(popcomp, school == .x)
      )
    } %>%
      broom::tidy(conf.int = TRUE, exponentiate = TRUE) %>%
      .[, c("term", "estimate", "conf.low", "conf.high")] %>%
      cbind(school = as.character(.x), .)
  ) %>% filter(term %in% "sex")

n <- popcomp %>% group_by(school) %>% 
  tally() %>% 
  mutate(school = as.character(school))

fits %>% left_join(n, by = "school") %>% 
  ggplot(aes(y = school, x = estimate, size = n)) +
  geom_vline(xintercept = 1, size = 1.5, color = "grey95") +
  geom_linerange(aes(y = school, xmin = conf.low, xmax = conf.high, size = 0.5), position = position_dodge(width = 0.5)) +
  geom_point(shape = 20, position = position_dodge(width = 0.5)) +
  labs(y = "Study", x = "", color = NULL) +
  theme_classic() +
  scale_size(guide = 'none')
```

@@

```{r}
popmis <- popmis[, c("school", "popular", "sex")] 
```

@@

This shows us that  `sex`  may be a useful imputation model predictor. Moreover, the missingness in `popular` may depend on the observed values of other variables.

<!-- Plot the histogram for teacher experience conditional on the missingness indicator of `popular`: -->

```{r pop-hist}
# ggmice(popmis, aes(sex)) +
#   geom_histogram(fill = "white") +
#   facet_grid(. ~ is.na(popular), scales = "free", labeller = label_both)

ggplot(popmis, aes(y = popular, group = sex)) +
  geom_boxplot() + 
  theme_classic()
```

<!-- This shows us that there are no apparent differences in the distribution of `texp` depending on the missingness indicator of `popular` (t = r t.test(popmis$texp ~ is.na(popmis$popular)) %>% broom::tidy(.) %>% .[, c("statistic", "p.value")] %>% round(., 3) %>% unlist() %>% paste(collapse = ",  p = ")). -->
<!--  [TODO: think about what is a meaningful rule of thumb to signal that the user should be worried?] -->

@@

Analyze the imputations:
```{r pop-ignored-fit, echo=TRUE, message=FALSE, warning=FALSE}
fit <- with(imp, 
            lmer(popular ~ 1 + sex  + (1 | school))) 
```

@@

Analyze the imputations:
```{r, echo=TRUE, message=FALSE, warning=FALSE}
fit <- with(imp, 
            lmer(popular ~ 1 + sex + (1 | school))) 
```

### 2022-06-23 Manuscript/Imputation_of_Incomplete_Multilevel_Data.tex
Add forest plot for popularity data

\begin{CodeChunk}
\begin{CodeInput}
R> mod <- popular ~ 1 + sex + texp + sex:texp + (1 + sex | school)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> popmis <- popmis[, c("school", "popular", "sex", "texp")] 
\end{CodeInput}
\end{CodeChunk}

@@

This shows us that both \texttt{sex} and \texttt{texp} may be useful
imputation model predictors. Moreover, the missingness in
\texttt{popular} may depend on the observed values of other variables.
We'll highlight one other variable to illustrate, but ideally one would
inspect all relations. The questions we'll ask are: `Does the missing
data of pupil popularity (\texttt{popular}) depend on observed teacher
popularity (\texttt{texp})?'. This can be evaluated statistically, but
visual inspection usually suffices. We'll make a histogram of
\texttt{texp} separately for the pupils with known popularity and
missing popularity.

Plot the histogram for teacher experience conditional on the missingness
indicator of \texttt{popular}:

\begin{CodeChunk}
\begin{CodeInput}
R> ggmice(popmis, aes(texp)) +
+   geom_histogram(fill = "white") +
+   facet_grid(. ~ is.na(popular), scales = "free", labeller = label_both)
\end{CodeInput}

@@

This shows us that there are no apparent differences in the distribution
of \texttt{texp} depending on the missingness indicator of
\texttt{popular} (t = -0.873, p = 0.383). {[}TODO: think about what is a
meaningful rule of thumb to signal that the user should be worried?{]}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> fit <- with(imp, 
+             lmer(popular ~ 1 + sex + texp + sex:texp + (1 + sex | school))) 
\end{CodeInput}
\end{CodeChunk}

@@

             Estimate Std.Error   t.value        df   P(>|t|)       RIV       FMI 
(Intercept)     3.436     0.179    19.207    11.230     0.000     1.480     0.653 
sex             1.210     0.208     5.812    13.898     0.000     1.157     0.591 
texp            0.101     0.014     7.261     7.456     0.000     2.738     0.784 
sex:texp       -0.020     0.017    -1.153     7.534     0.284     2.686     0.780

                            Estimate 
Intercept~~Intercept|school    0.158 
sex~~sex|school                0.212 
Intercept~~sex|school         -0.026 
Residual~~Residual             0.663 
ICC|school                     0.188

@@

             Estimate Std.Error   t.value        df   P(>|t|)       RIV       FMI 
(Intercept)     3.361     0.177    18.967    16.581     0.000     0.965     0.543 
sex             1.250     0.243     5.145     9.696     0.000     1.796     0.699 
texp            0.105     0.014     7.425     8.751     0.000     2.087     0.731 
sex:texp       -0.026     0.019    -1.388     6.889     0.208     3.202     0.810

                            Estimate 
Intercept~~Intercept|school    0.216 
sex~~sex|school                0.240 
Intercept~~sex|school         -0.045 
Residual~~Residual             0.615 
ICC|school                     0.259

@@

             Estimate Std.Error   t.value        df   P(>|t|)       RIV       FMI 
(Intercept)     3.333     0.177    18.832    16.459     0.000     0.972     0.545 
sex             1.356     0.193     7.014    37.019     0.000     0.490     0.362 
texp            0.106     0.012     8.870    13.221     0.000     1.222     0.606 
sex:texp       -0.034     0.012    -2.901    51.319     0.005     0.387     0.306

                            Estimate 
Intercept~~Intercept|school    0.213 
sex~~sex|school                0.307 
Intercept~~sex|school         -0.084 
Residual~~Residual             0.622 
ICC|school                     0.254

>>>>

\begin{CodeChunk}
\begin{CodeInput}
R> mod <- popular ~ 1 + sex + (1 | school)
\end{CodeInput}
\end{CodeChunk}

The true effect is:

\begin{center}\includegraphics{Imputation_of_Incomplete_Multilevel_Data_files/figure-latex/unnamed-chunk-3-1} \end{center}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> popmis <- popmis[, c("school", "popular", "sex")] 
\end{CodeInput}
\end{CodeChunk}

@@

This shows us that \texttt{sex} may be a useful imputation model
predictor. Moreover, the missingness in \texttt{popular} may depend on
the observed values of other variables.

\begin{CodeChunk}
\begin{CodeInput}
R> # ggmice(popmis, aes(sex)) +
R> #   geom_histogram(fill = "white") +
R> #   facet_grid(. ~ is.na(popular), scales = "free", labeller = label_both)
R> 
R> ggplot(popmis, aes(y = popular, group = sex)) +
+   geom_boxplot() + 
+   theme_classic()
\end{CodeInput}

@@

@@

\begin{CodeChunk}
\begin{CodeInput}
R> fit <- with(imp, 
+             lmer(popular ~ 1 + sex  + (1 | school))) 
\end{CodeInput}
\end{CodeChunk}

@@

             Estimate Std.Error   t.value        df   P(>|t|)       RIV       FMI 
(Intercept)     4.820     0.353    13.648     4.245     0.000    33.200     0.979 
sex             0.971     0.376     2.581     4.128     0.059    63.141     0.989

                            Estimate 
Intercept~~Intercept|school    0.259 
Residual~~Residual             1.065 
ICC|school                     0.197

@@

             Estimate Std.Error   t.value        df   P(>|t|)       RIV       FMI 
(Intercept)     5.121     0.390    13.125     4.242     0.000    33.503     0.979 
sex             0.673     0.418     1.611     4.121     0.180    66.343     0.989

                            Estimate 
Intercept~~Intercept|school    0.317 
Residual~~Residual             1.253 
ICC|school                     0.213

@@

             Estimate Std.Error   t.value        df   P(>|t|)       RIV       FMI 
(Intercept)     4.941     0.422    11.708     4.203     0.000    39.844     0.982 
sex             1.029     0.537     1.916     4.076     0.127   105.671     0.993

                            Estimate 
Intercept~~Intercept|school    0.306 
Residual~~Residual             1.308 
ICC|school                     0.195
